The Bolocam Galactic Plane Survey: Survey Description and Data Reduction

### 2009-04-18 methods/ms.tex
distmaps

\newcommand{\ammonia}{\mbox{{\rm NH}$_3$}}
\newcommand{\form}{H$_2$CO}
\newcommand{\td}{\mbox{$T_D$}}
\newcommand{\jj}[2]{\mbox{$J = #1\rightarrow#2$}}
\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{{\tt jaguirre@sas.upenn.edu}}
\newcommand{\lsim}{{_{<}\atop^{\sim}}}
\newcommand{\gsim}{{_{>}\atop^{\sim}}}
\newcommand{\etal}{{et al.\/}}
\newcommand{\ie}{{\em ie.\/}}
\newcommand{\cmq}{cm{$^{-3}$}}
\newcommand{\per}{$^{\rm{-1}}$}
\newcommand{\tc}{{$\theta^1$~Orionis~C}}
%\newcommand{\msol}{M{$_{\odot}$}}
\def\msol{\ifmmode {\>M_\odot}\else {$M_\odot$}\fi}
\newcommand{\lsol}{L{$_{\odot}$}}
\newcommand{\kms}{km~s{$^{-1}$}}
\newcommand{\hii}{H~{\sc ii}}
\newcommand{\Hii}{H~{\sc ii}}
\newcommand{\Ha}{\mbox{H$\alpha$}}
\newcommand{\sii}{S~{\sc ii}}
\newcommand{\Feii}{Fe~{\sc ii}}
\newcommand{\oi}{O~{\sc i}}
\newcommand{\nii}{N~{\sc ii}}
\newcommand{\oiii}{O~{\sc iii}}
\newcommand{\mgii}{Mg~{\sc ii}}
\newcommand{\tco}{{$^{13}$CO}}
\newcommand{\CO}{{$^{12}$CO}}
\newcommand{\Tco}{{$^{12}$CO}}
\newcommand{\co}{C{$^{18}$O}}
\newcommand{\Lsol}{L$_{\odot}$}
\newcommand{\Msol}{M$_{\odot}$}
%\newcommand{\C18o}{C$^{18}$O($1\rightarrow 0$)}
\newcommand{\Check}{{\bf ???}}
\newcommand{\mum}{\ensuremath{\mu \mathrm{m}}}
\newcommand{\flux}{flux density}
\newcommand{\solar}{\ensuremath{\odot}}
\newcommand{\lon}{\ensuremath{l}}

@@

We present the Bolocam Galactic Plane Survey (BGPS), a 1.1 mm continuum survey
of \bgpsarea\ square degrees of the Galactic Plane with 130 in the 1st quadrant,
31 in the 2nd, and 6 in the 3rd.  The BGPS represents
the first large area, systematic survey of the Northern Galactic Plane in the
millimeter continuum without pre-selected targets.  The survey has detected
\ncores\ clumps to a limiting non-uniform 1-$\sigma$ noise level in the range \bgpsdepthlow\ and
\bgpsdepthhigh\ mJy/beam RMS.  We present maps of 97 contiguous square degrees from
$\lon=-10.5$ to $\lon=86.5$ and $b=\pm0.5$ with a resolution of $\sim$35\arcsec, as
well as regions in IC1396, $\lon=111$ (NGC7538), W3/4/5, and Gem OB1.  The most
striking feature of the maps is the filamentary structure of millimeter emission
throughout the Galaxy.

@@

% TEST
% Useful, but not here
%\begin{figure}
%  \subfigure[$\lon=-10.5$ to $\lon=19.5$ Units are Jy/Beam.  The brightest
%	sources, e.g. Sgr B2, Sgr A, and sources near $\lon=10$ and
%	$\lon=13$, appear to be saturated, but this is only a display
%	artifact.  The noise is more pronounced from $\lon=-7$ to $\lon=-2$
%	because this region was observed less.]{\label{fig:BGPSMontage-a}\includegraphics[scale=0.75]{fig1_pg1_bw_inv_bar_label}}\\
%\pagebreak
%\subfigure[$\lon=19.5$ to $\lon=49.5$.  Units are Jy/Beam .  G34.3+0.15, W
%	51, W 43, W 49, and M 17 appear to be saturated, but this is
%	only a display artifact.]
%{\label{fig:BGPSMontage-b}\includegraphics[scale=0.75]{fig1_pg2_bw_inv_bar_label}}
%
%\end{figure}
%% --------------------------------
%

@@

\addtocounter{figure}{0}
%\begin{sidewaysfigure}[h]
\begin{figure}
    \label{fig:BGPSMontage}
  \begin{minipage}{6.5in}
    \begin{center}
      \includegraphics[scale=0.9]{fig1_pg1_bw_inv_bar_label}
      \caption{$\lon=-10.5$ to $\lon=19.5$ Units are Jy/Beam. 
%      The brightest
%	sources, e.g. Sgr B2, Sgr A, and sources near $\lon=10$ and
%	$\lon=13$, appear to be saturated, but this is only a display
%	artifact.
        The noise is more pronounced from $\lon=-7$ to $\lon=-2$
	because this region was observed less.}
    \end{center}
  \end{minipage}
\end{figure}

@@

\begin{figure}
  \begin{minipage}{6.5in} 
    \begin{center}
      \includegraphics[scale=0.9]{fig1_pg2_bw_inv_bar_label} 
      \caption{$\lon=19.5$ to $\lon=49.5$.  Units are Jy/Beam .  G34.3+0.15, W
	51, W 43, W 49, and M 17 appear to be saturated, but this is
	only a display artifact.  The $20 < \ell < 40$ region through the
        4-8 kpc molecular ring and approximately the termination of the galactic
        bar is particularly rich in clumps.}
    \end{center}
  \end{minipage}
\end{figure}

@@

The observations described here were acquired during six separate
observing sessions at the CSO over the course of two years.  The
observing epochs are given in Table \ref{tab:Observing}, along with
corresponding ranges for the zenith opacity $\tau_{225}$ of the CSO
tipper tau.  Because of the CSO weather multiplexing policy, Bolocam
observations were typically scheduled when $\tau_{225} > 0.06$.
Between each observing epoch, Bolocam was removed from its mount at
the re-imaged Cassegrain focus and stored warm.  Thus the flux
calibration and pointing model must be re-computed for each epoch to
allow for variations in the instrument and optics.  We found that the
flux calibration did not in fact differ significantly between epochs.
For the pointing model, we computed a well-constrained version for
Epoch V and aligned all images to this.

@@

Our basic observing strategy was to raster scan Bolocam by moving the
primary mirror of the CSO to modulate the astrophysical signal faster
than fluctuations in atmospheric opacity.  Each field was scanned with
alternating raster scans along $l$ and $b$.  Starting with Epoch III,
the fundamental scan unit was a $3\arcdeg \times 1\arcdeg$ block.
Each block was scanned with 23 scans along lines of constant $b$, and
67 scans along lines of constant $l$.  In both cases, the spacing
between adjacent rasters was 162\arcsec.  The total time for such a
scan was between 39 and 48 minutes.  The data were electronically
sampled at 10 Hz along the scan direction, slightly higher than the
Nyquist rate for the scan speed of 120\arcsec\ s$^{-1}$.

In order to fill in the instanteously undersampled focal plane, we
began using in Epoch II a field rotator to mechanically align the
focal plane at an optimal orientation along the scan direction to
improve the sampling orthogonal to the scan direction.  This is shown
in Figure \ref{fig:Array}.  Achieving good sampling is complicated by
a number of missing pixels.  A simulation was performed to determine
the optimal angle to rotate the array with respect to the scan
direction to account for the beam spacing and missing pixels.  During
the turnaround following a raster, the field rotator is adjusted to
this optimum angle.
%CHECK
% per Adam, 2009/3/26
Without the field rotator, the coverage shows variations of 100\% from
pixel to pixel in a single raster of a field.  With the rotator, this
is reduced to $\sim40\%$.

@@

The data were reduced with a custom pipeline written specifically for the BGPS.
A key element of the reduction is the iterative estimation of the
atmospheric fluctuations and the astrophysical signal.  The
atmospheric model is developed from a set of principal components of
the bolometer signal time series under the assumption that the bulk of
the correlated signal is atmospheric.  The iterative estimation
process then helps to restore correlated flux due to astrophysical
emission (Section \ref{sec:Mapping}).

The Bolocam observations are sensitive to structure between 35\arcsec\ and
about 200\arcsec\ (half the array field of view, see figure
\ref{fig:PCA_filter}).  An indicative analysis of the spatial filter function
of the iterative mapping scheme is given in \citet{enoch06}.  The data were
sampled onto a uniform 7\farcs2 grid.  Flux density calibrations were performed
using observations of Mars, Uranus, and Neptune, and they are accurate to
$\sim$20\%.  The flux calibration is described in section \ref{sec:FluxCalibration}.

%A calibration scheme capable of
%5\% systematic and $<$5\% statistical errors is being implemented as outlined in
%\citet{Sayers2009}.
%To account for offsets between the observations, each observation was mapped
%individually and aligned to a `master' field that has been tied to
%pointing calibrators. 
Noise properties across the survey are not completely uniform, but the
measured $1\sigma$ level is $\approx43\pm7$ mJy/beam within the
$b=\pm.5$ portion of the plane in the 1st quadrant.
%After the pipeline mapping, the fields were combined using the COMBINE
%command in IRAF with WCS offsets.

The absolute pointing accuracy of the survey has been measured to be
$\sim$6\arcsec\ RMS (figure \ref{fig:PointingModel}.  The Epoch V pointing model
was used to make pointing reference fields which were then median combined
using IRAF to make a higher S/N reference field with somewhat reduced pointing
errors (Section \ref{sec:Astrometry}).

@@

The absolute flux calibration is obtained from observations of Mars
and Uranus
% no good observations, per AG
%and Neptune 
(the ``primary calibrators'').  The millimeter-wave flux of these
planets is known to $\sim 5\%$ \citep{orton86,griffin93}.
Calibrations at (sub)millimeter wavelengths are strongly affected by
atmospheric opacity corrections.  Further, the detector responsivity
of Bolocam's bolometers is a non-linear function of the mean
atmospheric loading.

To relate observations of the primary calibrators to observations of
the BGPS fields, we make use of the following relation.  The
calibration ${\cal S}$, referenced to the detectors, is given by
\begin{equation}
\label{eq:Calibration}
{\cal S} \, \left[\mathrm{\frac{V}{Jy}}\right] = 
S(\tau) \eta A \exp{(-\tau)} \Delta\nu
\end{equation}
where $S$ is the bolometer responsivity ([V/W]), $\eta$ is the system
optical efficiency, $A$ is the effective telescope collecting area,
$\Delta \nu$ is the bandwidth, and $\tau$ is the line-of-sight,
in-band atmosphere opacity.  Under the assumption that the only power
variation on the detectors is due to the power from the atmosphere,
which may be parameterized by $\tau$, ${\cal S}$ is a single-valued
function of $\tau$.  We have used the mean bolometer resistance as a
proxy for $\tau$, since the bolometer resistance is a single-valued
function of loading.  {\it This quantity is monitored continuously for
all observations.}  Note that this calibration curve folds in both the
effects of changing atmospheric transmission and the changes in
the detector response with optical loading.

@@

A fit to the observed values for the primary calibrators was performed
for the each epoch separately.
%CHECK
The agreement between epochs was good, so a single combined
calibration was used for all data.
%We have found that the calibration can differ
%between successive cooldowns of the instrument by $\sim10\%$ \Check\
%but is generally stable over the period of order a month which
%constitutes each of the observing epochs.  
The calibration curve is shown in Figure \ref{fig:CalibrationCurves}.  The
resulting error on the calibration curve fit is less than 
%CHECK
$5\%$ (statistical) over the observed range of $\tau$.  A tally of all
contributions to the flux density error is given in Table
\ref{tab:ErrorBudget}.

The above flux density calibration only accounts for the average
calibration of bolometer Volts to Janskys.  It is also necessary to
account for the variation of bolometer response across the focal
plane.  We do this by monitoring the response to the atmosphere
emission in all bolometers.  Since the atmosphere is common to all
detectors, any changes in response are due to the intrinsic properties
of the detectors.  The change in relative response with loading for a
typical detector is shown in Figure \ref{fig:CalibrationCurves}.  The
actual application of the the method to the time series of the
bolometer data is shown in Figure \ref{fig:Flatfield}, and is
discussed further in Section \ref{sec:Mapping}.

@@

\subsection{Comparison to Other Surveys}
\label{sec:motte}

Comparison of the accuracy of the Bolocam flux density calibration
with the SCUBA 850 \mum\ results is complicated by the unknown effect
of the dust spectral index.  However, there are several large-area
surveys with the MAMBO instrument at 1.2 mm with which we can compare.
Figure \ref{fig:CygnusComparison} shows the comparison of fluxes
extracted from our map in the Cygnus Arm with those of
\citet{motte07} (M07).

The figures of the Cygnus-X region presented by \citet{motte07}, obtained at a
central wavelength of 1.2 mm using the 30 meter IRAM telescope, appear nearly
identical to the 1.1 mm Bolocam images.  The BGPS mapping strategy, larger
beam, and field of view provides better recovery of the large scale structure
of the dust continuum emission, but inferior angular resolution.  The MAMBO
filter is wider than the Bolocam filter and includes the CO 2-1 line.
Nevertheless, there is excellent agreement between the structure revealed by
the two data sets.

The Bolocam and M07 data sets have flux calibration consistent within about
20\%.  However, direct comparison with the fluxes listed in M07 is difficult
due to the factor of three difference in the size of the beams used.  For
example, the the brightest source in the Cyg-X region is DR 21.  While M07
quote a peak flux density per beam of 4.2 Jy/beam for the brightest feature in
the DR 21 complex (their peak N46) and 17.9 Jy/beam in a 0.19 x 0.14 pc box,
BGPS measures a peak flux of 17.5 Jy/beam (the beam being roughly 35\arcsec\ in
diameter).  Fluxes are best compared using sources that appear compact as
observed with the beam of the IRAM 30 meter telescope.

@@

\Table
{llllll}
{Comparison of \citet{motte07} IRAM measurements to BGPS measurements
in the Cygnus X region.  The IRAM data has been convolved to the BGPS beam
to make the flux density measurements.
%Note:  The \citet{motte07} integration box is given in terms of pc. 
%The assumed distance must be about 1.7 kpc \citep{schneider2006}    % no, it's 1.7kpc stated explicitly
}
{
%Source  &   BGPS           &  Motte \\
%        &   peak mJy/beam  &  mJy in box (see Motte+ Table 1)
Motte Source &  Motte Table 1   &  Motte scaled  & BGPS  &  Comments \\
      Number &  integrated flux &  Jy/bm         & Jy/bm &           \\
}
{tab:CygnusComparison}
{
3  &    1.59 [0.11x0.09] & 2.61      &  2.30  &  DR 17                               \\
11 &    0.34 [0.11x0.06] & 0.48      &  0.50  &  West SW of DR21                     \\
12 &    1.63 [0.11x0.09] & 2.60      &  2.20  &    "                                 \\
16 &    0.55 [0.12x0.12] & 0.82      &  0.72  &    "                                 \\
26 &    0.34 [0.20x0.10] & 0.48      &  0.44  &    "                                 \\
29 &    0.54 [0.25x0.13] & 1.10      &  0.62  &    "                                 \\
30 &   10.54 [0.15x0.10] &13.5       &  9.80  &  W75(N)                              \\
44 &    8.43 [0.14x0.11] &14.5       & 12.7   &  DR21 (OH)                           \\
46 &   17.90 [0.08x0.03] &18.4       & 15.8   &  DR21 f-f, maser (H2O)               \\
48 &    3.72 [0.17x0.11] & 9.4       &  7.8   &  DR21 (OH)-S                         \\
51 &    2.37 [0.24x0.03] & 5.5       &  4.1   &  DR21 FIR2                           \\
53 &    1.60 [0.14x0.09] & 4.3       &  2.7   &  North of DR21  FIR3                 \\
56 &    0.60 [0.12x0.08] & 0.81      &  0.75  &  SE of DR21 47                       \\
58 &    2.73 [0.23x0.19] & 2.40      &  1.64  &  IRAS 20375+4109                     \\
59 &    0.37 [0.12x0.05] & 1.3       &  0.93  &  East of IRAS 20375+4109             \\
60 &    0.56 [0.16x0.13] & 1.00      &  0.78  &  East of IRAS 20375+4109             \\
61 &    0.23 [0.12x0.09] & 0.60      &  0.49  &  MSX 81.5168+0.01926                 \\
62 &    0.50 [0.18x0.12] & 1.23      &  0.80  &  between 63 \& IRAS 20375+4109       \\
63 &    1.09 [0.08x0.05] & 1.60      &  1.50  &  DR22                                \\
65 &    0.78 [0.16x0.12] & 1.15      &  1.10  &                                      \\
66 &    0.25 [0.15x0.06] & 0.63      &  0.76  &                                      \\
68 &    0.84 [0.10x0.10] & 1.01      &  0.94  &                                      \\
69 &    1.28 [0.47x0.18] & 1.12      &  1.20  &                                      \\
72 &    0.54 [0.28x0.13] & 0.58      &  0.47  &                                      \\
}

\FigureTwo{fluxcal_errorbars}{calib_vs_dc}
%{rel_flux_cal_050303_ob3_to_050405_o66}
%{calibration_plot_2004}
{Left: Average calibration curve [V/Jy] versus the mean detector voltage, a
proxy for atmospheric loading.  Red points are observations of Mars and black
of Uranus.  The black line is a 2nd-order fit with 0,0 forced (no response if no
measurable potential difference) .  Right: Scaling of the relative
response to the atmosphere for a single detector compared to the array
median.}{fig:CalibrationCurves}{1.0}

>>>>

\newcommand{\ammonia}{\mbox{{\rm NH}$_3$}}
\newcommand{\form}{H$_2$CO}
\newcommand{\td}{\mbox{$T_D$}}
\newcommand{\jj}[2]{\mbox{$J = #1\rightarrow#2$}}
\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{{\tt jaguirre@sas.upenn.edu}}
\newcommand{\lsim}{{_{<}\atop^{\sim}}}
\newcommand{\gsim}{{_{>}\atop^{\sim}}}
\newcommand{\etal}{{et al.\/}}
\newcommand{\ie}{{\em ie.\/}}
\newcommand{\cmq}{cm{$^{-3}$}}
\newcommand{\per}{$^{\rm{-1}}$}
\newcommand{\tc}{{$\theta^1$~Orionis~C}}
%\newcommand{\msol}{M{$_{\odot}$}}
\def\msol{\ifmmode {\>M_\odot}\else {$M_\odot$}\fi}
\newcommand{\lsol}{L{$_{\odot}$}}
\newcommand{\kms}{km~s{$^{-1}$}}
\newcommand{\hii}{H~{\sc ii}}
\newcommand{\Hii}{H~{\sc ii}}
\newcommand{\Ha}{\mbox{H$\alpha$}}
\newcommand{\sii}{S~{\sc ii}}
\newcommand{\Feii}{Fe~{\sc ii}}
\newcommand{\oi}{O~{\sc i}}
\newcommand{\nii}{N~{\sc ii}}
\newcommand{\oiii}{O~{\sc iii}}
\newcommand{\mgii}{Mg~{\sc ii}}
\newcommand{\tco}{{$^{13}$CO}}
\newcommand{\CO}{{$^{12}$CO}}
\newcommand{\Tco}{{$^{12}$CO}}
\newcommand{\co}{C{$^{18}$O}}
\newcommand{\Lsol}{L$_{\odot}$}
\newcommand{\Msol}{M$_{\odot}$}
%\newcommand{\C18o}{C$^{18}$O($1\rightarrow 0$)}
\newcommand{\Check}{{\bf ???}}
\newcommand{\mum}{\ensuremath{\mu \mathrm{m}}}
\newcommand{\flux}{flux density}
\newcommand{\solar}{\ensuremath{\odot}}
\newcommand{\lon}{\ensuremath{l}}
\newcommand\jyb{Jy~beam$^{-1}$}

@@

We present the Bolocam Galactic Plane Survey (BGPS), a 1.1 mm continuum survey
of \bgpsarea\ square degrees of the Galactic Plane with 130 in the 1st quadrant,
31 in the 2nd, and 6 in the 3rd.  The BGPS represents
the first large area, systematic survey of the Northern Galactic Plane in the
millimeter continuum without pre-selected targets.  The survey has detected
\ncores\ clumps to a limiting non-uniform 1-$\sigma$ noise level in the range \bgpsdepthlow\ and
\bgpsdepthhigh\ m\jyb RMS.  We present maps of 97 contiguous square degrees from
$\lon=-10.5$ to $\lon=86.5$ and $b=\pm0.5$ with a resolution of $\sim$35\arcsec, as
well as regions in IC1396, $\lon=111$ (NGC7538), W3/4/5, and Gem OB1.  The most
striking feature of the maps is the filamentary structure of millimeter emission
throughout the Galaxy.

@@

% TEST
% Useful, but not here
%\begin{figure}
%  \subfigure[$\lon=-10.5$ to $\lon=19.5$ Units are \jyb.  The brightest
%	sources, e.g. Sgr B2, Sgr A, and sources near $\lon=10$ and
%	$\lon=13$, appear to be saturated, but this is only a display
%	artifact.  The noise is more pronounced from $\lon=-7$ to $\lon=-2$
%	because this region was observed less.]{\label{fig:BGPSMontage-a}\includegraphics[scale=0.75]{fig1_pg1_bw_inv_bar_label}}\\
%\pagebreak
%\subfigure[$\lon=19.5$ to $\lon=49.5$.  Units are \jyb .  G34.3+0.15, W
%	51, W 43, W 49, and M 17 appear to be saturated, but this is
%	only a display artifact.]
%{\label{fig:BGPSMontage-b}\includegraphics[scale=0.75]{fig1_pg2_bw_inv_bar_label}}
%
%\end{figure}
%% --------------------------------
%

@@

\addtocounter{figure}{0}
%\begin{sidewaysfigure}[h]
\begin{figure}
    \label{fig:BGPSMontage}
  \begin{minipage}{6.5in}
    \begin{center}
      \includegraphics[scale=0.9]{fig1_pg1_bw_inv_bar_label}
      \caption{$\lon=-10.5$ to $\lon=19.5$ Units are \jyb. 
%      The brightest
%	sources, e.g. Sgr B2, Sgr A, and sources near $\lon=10$ and
%	$\lon=13$, appear to be saturated, but this is only a display
%	artifact.
        The noise is more pronounced from $\lon=-7$ to $\lon=-2$
	because this region was observed less.}
    \end{center}
  \end{minipage}
\end{figure}

@@

\begin{figure}
  \begin{minipage}{6.5in} 
    \begin{center}
      \includegraphics[scale=0.9]{fig1_pg2_bw_inv_bar_label} 
      \caption{$\lon=19.5$ to $\lon=49.5$.  Units are \jyb .  G34.3+0.15, W
	51, W 43, W 49, and M 17 appear to be saturated, but this is
	only a display artifact.  The $20 < \ell < 40$ region through the
        4-8 kpc molecular ring and approximately the termination of the galactic
        bar is particularly rich in clumps.}
    \end{center}
  \end{minipage}
\end{figure}

@@

The maps presented here were acquired during six separate
observing sessions at the CSO over the course of two years.  The
observing epochs are given in Table \ref{tab:Observing}, along with
corresponding ranges for the zenith opacity $\tau_{225}$ of the CSO
tipper tau.  Bolocam
observations were typically scheduled when $\tau_{225} > 0.06$.
Between observing epochs, Bolocam was removed from its mount at
the re-imaged Cassegrain focus and stored warm.  Thus the flux
calibration and pointing model must be re-computed for each epoch to
allow for variations in the instrument and optics.  We found that the
flux calibration did not in fact differ significantly between epochs.
We observed additional pointing sources during Epoch V and therefore
used it as the master pointing model.

@@

[Reword section?] Our basic observing strategy was to raster scan Bolocam by moving the
primary mirror of the CSO to modulate the astrophysical signal faster
than fluctuations in atmospheric opacity.  Each field was scanned with
alternating raster scans along $l$ and $b$.  In Epoch I-II, the fundamental
scan unit was a $1\arcdeg \times 1\arcdeg$ block, scanned 23 times in \lon and
$b$.  Starting with Epoch III, the fundamental scan unit was changed to a
$3\arcdeg \times 1\arcdeg$ block for increased mapping efficiency.  Each block
was scanned with 23 scans along lines of constant $b$, and 67 scans along lines
of constant $l$.  In both cases, the spacing between adjacent rasters was
162\arcsec.  The total time for such a scan was between 39 and 48 minutes.  The
data were electronically sampled at 10 Hz along the scan direction, slightly
higher than the Nyquist rate for the scan speed of 120\arcsec\ s$^{-1}$.

In order to fill in the instanteously undersampled focal plane, we
began using in Epoch II a field rotator to mechanically align the
focal plane at an optimal orientation along the scan direction to
improve the sampling orthogonal to the scan direction.  This is shown
in Figure \ref{fig:Array}.  Achieving good sampling is complicated by
a number of missing bolometers.  A simulation was performed to determine
the optimal angle to rotate the array with respect to the scan
direction to account for the beam spacing and missing bolometers.  During
the turnaround following a raster, the field rotator is adjusted to
this optimum angle.
%CHECK
% per Adam, 2009/3/26
Without the field rotator, the coverage shows variations of 100\% from
pixel to pixel in a single raster of a field.  With the rotator, this
is reduced to $\sim40\%$.

@@

The data were reduced with a custom pipeline written specifically for the BGPS.
A key element of the reduction is the iterative estimation of the
atmospheric fluctuations and the astrophysical signal.  The
atmospheric model is developed from a set of principal components of
the bolometer signal time series under the assumption that the bulk of the
correlated signal is atmospheric (under this assumption, astrophysical extended
structures are treated as atmospheric noise).  The iterative estimation
process then helps to restore correlated flux due to astrophysical
emission (Section \ref{sec:Mapping}).

The Bolocam observations are sensitive to structure between 35\arcsec\ and
about 200\arcsec\ (half the array field of view, see figure
\ref{fig:PCA_filter}).  An indicative analysis of the spatial filter function
of the iterative mapping scheme is given in \citet{enoch06}, and we present a detailed analysis
of the filter function in our survey in figure \ref{fig:PCA_filter}.  The data were
sampled onto a uniform 7\farcs2 grid.  Flux density calibrations were performed
using observations of Mars, Uranus, and Neptune, and they are accurate to
$\sim$20\%.  The flux calibration is described in section \ref{sec:FluxCalibration}.

%A calibration scheme capable of
%5\% systematic and $<$5\% statistical errors is being implemented as outlined in
%\citet{Sayers2009}.
%To account for offsets between the observations, each observation was mapped
%individually and aligned to a `master' field that has been tied to
%pointing calibrators. 
Noise properties across the survey are not completely uniform, but the
measured $1\sigma$ level is $\approx43\pm7$ m\jyb within the
$b=\pm.5$ portion of the plane in the 1st quadrant.
%After the pipeline mapping, the fields were combined using the COMBINE
%command in IRAF with WCS offsets.

A separate model was constructed for each epoch, but epoch V was used as the
primary pointing calibration and is shown in figure \ref{fig:PointingModel}.
In epoch V and VI, pointing sources were observed more frequently to obtain a
better-constrained master pointing model. The absolute pointing accuracy was
$\sim6\arcsec$ RMS.  The Epoch V pointing model was used to make pointing
reference fields which were then median combined using IRAF to make a higher
S/N reference field with somewhat reduced pointing errors (Section
\ref{sec:Astrometry}).

@@

The absolute flux calibration is derived from observations of Mars
and Uranus
% no good observations, per AG
%and Neptune 
(the ``primary calibrators'').  The millimeter-wave flux of these
planets is known to $\sim 5\%$ \citep{orton86,griffin93}.
Calibrations at (sub)millimeter wavelengths are strongly affected by
atmospheric opacity corrections.  Further, the detector responsivity
of Bolocam's bolometers is a non-linear function of the mean
atmospheric loading.

To relate observations of the primary calibrators to observations of
the BGPS fields, we make use of the following relation.  The
calibration ${\cal S}$, referenced to the detectors, is given by
\begin{equation}
\label{eq:Calibration}
{\cal S} \, \left[\mathrm{\frac{V}{Jy}}\right] = 
S(\tau) \eta A \exp{(-\tau)} \Delta\nu
\end{equation}
where $S$ is the bolometer responsivity ([V/W]), $\eta$ is the system
optical efficiency, $A$ is the effective telescope collecting area,
$\Delta \nu$ is the bandwidth, and $\tau$ is the line-of-sight,
in-band atmosphere opacity.  Under the assumption that the only power
variation on the detectors is due to the power from the atmosphere (i.e.
that astronomical sources are faint relative to the atmosphere),
which may be parameterized by $\tau$, ${\cal S}$ is a single-valued
function of $\tau$.  We have used the mean bolometer resistance as a
proxy for $\tau$, since the bolometer resistance is a single-valued
function of loading.  {\it This quantity is monitored continuously for
all observations.}  Note that this calibration curve folds in both the
effects of changing atmospheric transmission and the changes in
the detector response with optical loading.

@@

A fit to the observed values (measured potential difference vs. known flux
density) for the primary calibrators was performed for the each epoch
separately.
%CHECK
The agreement between epochs was good, so a single combined
calibration was used for all data.
%We have found that the calibration can differ
%between successive cooldowns of the instrument by $\sim10\%$ \Check\
%but is generally stable over the period of order a month which
%constitutes each of the observing epochs.  
The calibration curve is shown in Figure \ref{fig:CalibrationCurves}.  The
resulting error on the calibration curve fit is less than 
%CHECK
$5\%$ (statistical) over the observed range of $\tau$.  A tally of all
contributions to the flux density error is given in Table
\ref{tab:ErrorBudget}.

The above flux density calibration only accounts for the average
calibration of bolometer Volts to Janskys.  It is also necessary to
account for the variation of bolometer response across the focal
plane.  We do this by monitoring the response to the atmosphere
emission in all bolometers (astrophysical signal is negligible except at SgrB2).  Since the
atmosphere is common to all detectors [Anyone - why is the atmosphere common across 7.5'
at the mm?  Brief discussion of atmosphere physics would be helpful], any changes in response are due to the
intrinsic properties of the detectors.  The change in relative response with
loading for a typical detector is shown in Figure \ref{fig:CalibrationCurves}.
The actual application of the method to the time series of the bolometer
data is shown in Figure \ref{fig:Flatfield}, and is discussed further in
Section \ref{sec:Mapping}.

[move this?] In figure \ref{fig:Flatfield}, each step of the iterative cleaning process is illustrated.

@@

\subsection{Comparison to Other Surveys}
\label{sec:motte}
\bf{Claudia's comments:}
In general, this section could use some reorganizing such that it sets out why
this comparison field was chosen first (best available data?), describes any
relevant differences in the datasets (eg filter bandpasses, CO possibly in
MAMBO filter), describes how the comparison was done, and then represents the
results.  It seems out of order at the moment.  A reference for CO
contamination causing detectable contamination (at 850um, not 1mm) is
Hatchell, J., \& Dunham, M.~M.\ 2009, arXiv:0904.1163

Comparison of the BGPS flux density calibration
with SCUBA 850 \mum\ results is complicated by the unknown effect
of the dust spectral index.  However, there are several large-area
surveys with the MAMBO instrument at 1.2 mm with which we can compare.
Figure \ref{fig:CygnusComparison} shows the comparison of fluxes
extracted from our map in the Cygnus Arm with those of
\citet{motte07} (M07).

The figures of the Cygnus-X region presented by \citet{motte07}, obtained at a
central wavelength of 1.2 mm using the 30 meter IRAM telescope, appear nearly
identical to the 1.1 mm Bolocam images.  The MAMBO filter is wider than the
Bolocam filter and includes the CO 2-1 line.  Nevertheless, there is excellent
qualitative agreement between the structure revealed by the two data sets on size
scales to which both are sensitive.

[John - may want to remove N46 and other extended sources from comparison.  This 
paragraph needs cleanup]
Direct comparison with the fluxes listed in M07 is difficult
due to the factor of three difference in the size of the beams used.  For
example, the the brightest source in the Cyg-X region is DR 21.  While M07
quote a peak flux density per beam of 4.2 \jyb for the brightest feature in
the DR 21 complex (their peak N46) and 17.9 Jy in a 0.19 x 0.14 pc box,
BGPS measures a peak flux of 17.5 \jyb (the beam being roughly 35\arcsec\ in
diameter).  Fluxes are best compared using sources that appear compact as
observed with the beam of the IRAM 30 meter telescope. Although we list comparisons
for non-compact sources, our flux calibratio comparison was made using compact sources.
The Bolocam and M07 data sets have flux calibration consistent within about
20\%.

@@

\Table
{lllllll}
{Comparison of \citet{motte07} IRAM measurements to BGPS measurements
in the Cygnus X region.  The IRAM data has been convolved to the BGPS beam
to make the flux density measurements. [John - what does `scaled' mean?]
%Note:  The \citet{motte07} integration box is given in terms of pc. 
%The assumed distance must be about 1.7 kpc \citep{schneider2006}    % no, it's 1.7kpc stated explicitly
}
{
%Source  &   BGPS           &  Motte \\
%        &   peak m\jyb  &  mJy in box (see Motte+ Table 1)
Motte Source &  Motte Table 1   &  Motte scaled  & BGPS  & Ratio &  Comments \\
      Number &  integrated flux &  \jyb          & \jyb  &       &           \\
}
{tab:CygnusComparison}
{
3  &    1.59 [0.11x0.09] & 2.61      &  2.30  & & DR 17                               \\
11 &    0.34 [0.11x0.06] & 0.48      &  0.50  & & West SW of DR21                     \\
12 &    1.63 [0.11x0.09] & 2.60      &  2.20  & &   "                                 \\
16 &    0.55 [0.12x0.12] & 0.82      &  0.72  & &   "                                 \\
26 &    0.34 [0.20x0.10] & 0.48      &  0.44  & &   "                                 \\
29 &    0.54 [0.25x0.13] & 1.10      &  0.62  & &   "                                 \\
30 &   10.54 [0.15x0.10] &13.5       &  9.80  & & W75(N)                              \\
44 &    8.43 [0.14x0.11] &14.5       & 12.7   & & DR21 (OH)                           \\
46 &   17.90 [0.08x0.03] &18.4       & 15.8   & & DR21 f-f, maser (H2O)               \\
48 &    3.72 [0.17x0.11] & 9.4       &  7.8   & & DR21 (OH)-S                         \\
51 &    2.37 [0.24x0.03] & 5.5       &  4.1   & & DR21 FIR2                           \\
53 &    1.60 [0.14x0.09] & 4.3       &  2.7   & & North of DR21  FIR3                 \\
56 &    0.60 [0.12x0.08] & 0.81      &  0.75  & & SE of DR21 47                       \\
58 &    2.73 [0.23x0.19] & 2.40      &  1.64  & & IRAS 20375+4109                     \\
59 &    0.37 [0.12x0.05] & 1.3       &  0.93  & & East of IRAS 20375+4109             \\
60 &    0.56 [0.16x0.13] & 1.00      &  0.78  & & East of IRAS 20375+4109             \\
61 &    0.23 [0.12x0.09] & 0.60      &  0.49  & & MSX 81.5168+0.01926                 \\
62 &    0.50 [0.18x0.12] & 1.23      &  0.80  & & between 63 \& IRAS 20375+4109       \\
63 &    1.09 [0.08x0.05] & 1.60      &  1.50  & & DR22                                \\
65 &    0.78 [0.16x0.12] & 1.15      &  1.10  & &                                     \\
66 &    0.25 [0.15x0.06] & 0.63      &  0.76  & &                                     \\
68 &    0.84 [0.10x0.10] & 1.01      &  0.94  & &                                     \\
69 &    1.28 [0.47x0.18] & 1.12      &  1.20  & &                                     \\
72 &    0.54 [0.28x0.13] & 0.58      &  0.47  & &                                     \\
}
